---
title: "Quiz 01: Data wrangling"
Date: "Wenesday, March 15"
author: "Aicha Sidiya"
date: "3/15/2023"
editor_options: 
  chunk_output_type: console
output:
  pdf_document: default
  html_document: default
---

# Packages

```{r load-packages, message = FALSE}
library(tidyverse)
```

# Data

```{r load-data, message = FALSE}
lego <- read_csv("data/lego.csv")
```

# Exercise 1

```{r}
lego <- filter(lego, !is.na(retail_price), !is.na(pieces), pieces != 0)
```
Removing these sets shortned the dataframe but made it more expressive
# Exercise 2

```{r}
lego2 <- arrange(lego, desc(retail_price))
lego2[1:3, ] #print first 3 rows 
```

The three most expensive sets are Millennium Falcon which costs 800\$ and has 7541 pieces, the second most expensive set is Connections Kit which costs 755\$ and has 2455 pices, the third most expensive set is Death Star whcih costs 500\$ and contains 4016 pieces. 

# Exercise 3

```{r}
lego <- mutate(lego, price_per_piece=retail_price/pieces)
```

# Exercise 4

```{r}
lego4 <- arrange(lego, desc(price_per_piece))
select(lego4[1:5, ], name, themegroup, theme, pieces, price_per_piece)
```

The sets with the highest prices are the sets consisting of 1 piece only. 

# Exercise 5

```{r}
lego5 <- filter(lego, theme == "The Lord of the Rings")
lego6 <- group_by(lego5, subtheme)
summarize(lego6, cheap_lego = min(retail_price),
          expensive_lego = max(retail_price))
```

# Exercise 6

```{r}
lego7 <- group_by(lego, year)
yearly_trends <- summarize(lego7, setsNumber = n(), price_per_piece_median = median(price_per_piece))
```

# Exercise 7

```{r}
library(ggplot2)
ggplot(data=yearly_trends, mapping=aes(x=price_per_piece_median, y=year, size = setsNumber)) + geom_point() + labs(title = "Yearly trend of price per piece according to number of sets") 
```

The sisze of the sets increse over the years.
